---
title: "An Easy Way to Install R Packages on GitHub"
author: Koji MAKIYAMA (@hoxo_m)
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

[![Travis-CI Build Status](https://travis-ci.org/hoxo-m/githubinstall.svg?branch=master)](https://travis-ci.org/hoxo-m/githubinstall)
[![CRAN Version](http://www.r-pkg.org/badges/version/githubinstall)](http://cran.rstudio.com/web/packages/githubinstall)
[![Coverage Status](https://coveralls.io/repos/github/hoxo-m/githubinstall/badge.svg?branch=master)](https://coveralls.io/github/hoxo-m/githubinstall?branch=master)

## 1. Overview

A growing number of R packages are created by various people in the world.
A part of the cause of the growth is **devtools** package that makes it easy to develop R packages.
The **devtools** package not only facilitates the process to develop R packages but also provides another way to distribute R packages.

To distribute R packages, the CRAN is commonly used.
You can install the packages are available on CRAN as follows.

```{r eval=FALSE}
install.packages("dplyr")
```

The **devtools** package provides `install_github()` that enables installing packages on GitHub.

```{r eval=FALSE}
library(devtools)
install_github("hadley/dplyr")
```

Therefore, developers can distribute R packages that is developping on GitHub.
For example, Twitter Inc. provides **AnomalyDetection** package on GitHub but it is not available on CRAN.[^1]

[^1]:[Introducing practical and robust anomaly detection in a time series](https://blog.twitter.com/2015/introducing-practical-and-robust-anomaly-detection-in-a-time-series)

The **githubinstall** package provides an easy way to install packages on GitHub.

```{r eval=FALSE}
library(githubinstall)
githubinstall("AnomalyDetection")
```

In this case, you will succeed to install **AnomalyDetection** package as follows.

```{r eval=FALSE}
githubinstall("anomaly detection")
```

## 2. Installation

You can install the package from GitHub.

```{r eval=FALSE}
install.packages("devtools") # if you have not installed "devtools" package
devtools::install_github("hoxo-m/githubinstall")
```

The source code for **githubinstall** package is available on GitHub at

- https://github.com/hoxo-m/githubinstall.

## 3. Details

This package completes `username` automatically.

```{r}
library(githubinstall)
```

```{r eval=FALSE}
githubinstall("AnomalyDetection")
# This is same as devtools::install_github("twitter/AnomalyDetection")
```

Or

```{r eval=FALSE}
gh_install_packages("AnomalyDetection", build_vignettes = F)
```

```
Downloading GitHub repo hadley/multidplyr@master
Installing multidplyr
...
* DONE (multidplyr)
```

You can show the list of repositories on GitHub by `username`.

```{r}
hadleyverse <-  gh_get_package_info("hadley")
head(hadleyverse)
```

You can guess repository names or user names.

```{r}
gh_guess_username("hadly")
```

## 4. Related Work

- devtools: [Tools to make an R developer's life easier](https://github.com/hadley/devtools)
- [ghit: Lightweight GitHub Package Installer](https://github.com/cloudyr/ghit)
- drat: [Drat R Archive Template](https://github.com/eddelbuettel/drat)
- pacman: [A package management tools for R](https://github.com/trinker/pacman)
- remotes: [Install R packages from GitHub, Bitbucket, git, svn repositories, URLs](https://github.com/MangoTheCat/remotes)
