---
title: "githubinstall: A Helpful Way to Install R Packages Hosted on GitHub"
author: "Koji MAKIYAMA (@hoxo_m)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{githubinstall: A Helpful Way to Install R Packages Hosted on GitHub}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Overview

A growing number of R packages are created by various people in the world.
A part of the cause of it is the **devtools** package that makes it easy to develop R packages [[1]](https://www.rstudio.com/products/rpackages/devtools/).
The **devtools** package not only facilitates the process to develop R packages but also provides an another way to distribute R packages.

When developers publish R packages that created by them, the CRAN [[2]](https://cran.r-project.org) is commonly used.
You can install packages that are available on CRAN using `install.package()`.
For example, you can install **dplyr** package as follows:

```{r eval=FALSE}
install.packages("dplyr")
```

The **devtools** package provides `install_github()` that enables installing packages from GitHub.

```{r eval=FALSE}
library(devtools)
install_github("hadley/dplyr")
```

Therefore, developers can distribute R packages that are developing on GitHub.
In addition, there are some developers that they have no intention to submit to CRAN.
For instance, Twitter, Inc. provides **AnomalyDetection** package on GitHub but they won't submit to CRAN [[3]](https://blog.twitter.com/2015/introducing-practical-and-robust-anomaly-detection-in-a-time-series).
You can install such packages easily using **devtools**.

```{r eval=FALSE}
library(devtools)
install_github("twitter/AnomalyDetection")
```

There is a difference between `install.packages()` and `install_github()` in the required argument.
`install.packages()` takes package names, while `install_github()` needs repository names.
It means that when you want to install a package on GitHub you must remember its repository name correctly.

The trouble is that the usernames of GitHub are often hard to remember.
Developers consider their package names so that users can understand their functionalities intuitively.
However, they often decide username incautiously.
For instance, **ggfortify** is a great package on GitHub, but who created it?
What is its username?
The answer is *sinhrks* [[4]](https://github.com/sinhrks/ggfortify).
It seems to be difficult to remember it.

The **githubinstall** package provides a way to install packages on GitHub by only their package names just like `install.packages()`.

```{r eval=FALSE}
library(githubinstall)
githubinstall("AnomalyDetection")
```

```
Suggestion:
 - twitter/AnomalyDetection  Anomaly Detection with R
Do you want to install the package (Y/n)?  
```

`githubinstall()` suggests GitHub repositories from input package names and asks whether you really install it.

Furthermore, you may succeed in installing packages from a faint memory because our package automatically corrects its spelling by fuzzy string search.

```{r eval=FALSE}
githubinstall("AnomaryDetection")
githubinstall("AnomalyDetect")
githubinstall("anomaly-detection")
```

## 2. Installation

You can install the **githubinstall** package from CRAN.

```{r eval=FALSE}
install.packages("githubinstall")
```

You can also install the package from GitHub.

```{r eval=FALSE}
install.packages("devtools") # if you have not installed "devtools" package
devtools::install_github("hoxo-m/githubinstall")
```

The source code for **githubinstall** package is available on GitHub at

- https://github.com/hoxo-m/githubinstall.

## 3. Details

The **githubinstall** package provides several useful functions.

- `gh_install_packages()` or `githubinstall()`
- `gh_suggest()`
- `gh_suggest_username()`
- `gh_list_packages()`
- `gh_search_packages()`
- `gh_show_source()`
- `gh_update_package_list()`

The functions have common prefix `gh`.
`githubinstall()` is an alias of `gh_install_packages()`.

To use these functions, first, you should load the package as follows.

```{r}
library(githubinstall)
```

### 3.1. Install Packages from GitHub

`gh_install_packages()` enables to install packages on GitHub by only package names.

```{r eval=FALSE}
gh_install_packages("AnomalyDetection")
```

```
Suggestion:
 - twitter/AnomalyDetection  Anomaly Detection with R
Do you want to install the package (Y/n)?  
```

The function suggests GitHub repositories.
If you type 'Y' or 'y' and press 'Enter' (the default is 'Y'), then installation of the package will begin.
The suggestion is made of looking for a list of R packages on GitHub.
The list is provided by [Gepuro Task Views](http://rpkg.gepuro.net).

If multiple candidates are found, you can select one of them.

```{r eval=FALSE}
gh_install_packages("cats")
```

```
Select a number or, hit 0 to cancel. 

1: amurali2/cats      cats
2: danielwilhelm/cats No description or website provided.
3: hilaryparker/cats  An R package for cat-related functions #rcatladies
4: lolibear/cats      No description or website provided.
5: rafalszota/cats    No description or website provided.
6: tahir275/cats      ff

Selection: 
```

`githubinstall()` is an alias of `gh_install_packages()`.

```{r eval=FALSE}
githubinstall("AnomalyDetection")
```

#### 3.1.1. Specify Git References (Branch, Tag, Commit and Pull Request)

**A new feature has added.**

You can install packages with specifying Git references (branch, tag, commit and pull request).

Developers are divided in policy to manage R packages on GitHub.
If a package is going to be developed in "develop" branch, you may want to install the package from the branch.

`gh_install_packages()` has `ref` argument to specify Git references. 
For instance, you can install **awaptools** from the ["develop" branch](https://github.com/swish-climate-impact-assessment/awaptools/tree/develop) as follows:

```{r eval=FALSE}
gh_install_packages("awaptools", ref = "develop")
```

You may sometimes encounter failing to install packages because its repository HEAD is broken.
In such case, you can specify a tag or commit to `ref`.
In almost cases, tags are added on an unbroken commit.
For instance, you can install **densratio** from the ["v0.0.3" tag](https://github.com/hoxo-m/densratio/releases/tag/v0.0.3) as follows:

```{r eval=FALSE}
gh_install_packages("densratio", ref = "v0.0.3")
```

If you can find no such tags, the previous commit may be not broken.
For instance, you can install **densratio** from the ["e8233e6" commit](https://github.com/hoxo-m/densratio/commit/e8233e651dbef2b34a8c9c2e4432594a13ea8de7) as follows:

```{r eval=FALSE}
gh_install_packages("densratio", ref = "e8233e6")
```

Finally, you may find a patch for fixing bugs as a pull request.
In such case, you can specify pull requests to `ref` using `github_pull()`.
For instance, you can install **dplyr** from the [pull request #2058](https://github.com/hadley/dplyr/pull/2058) as follows:

```{r eval=FALSE}
gh_install_packages("dplyr", ref = github_pull("#2058"))
```

### 3.2. Suggest Repositories

`gh_install_packages()` prompts you to install the suggested packages.
But you may just want to know what will be suggestions.

`gh_suggest()` returns the suggested repository names as a vector.

```{r}
gh_suggest("AnomalyDetection")
```

```{r}
gh_suggest("cats")
```

In addition, `gh_suggest_username()` is useful when you want to know usernames from a faint memory.

```{r}
gh_suggest_username("hadly")
```

```{r}
gh_suggest_username("yuhui")
```

### 3.3. List the Packages

`gh_list_packages()` returns a list of R package repositories on GitHub as `data.frame`.

For example, if you want to get the repositories that have been created by *hadley*, run the following.

```{r eval=FALSE}
hadleyverse <- gh_list_packages(username = "hadley")
head(hadleyverse)
```

```{r echo=FALSE}
hadleyverse <- gh_list_packages(username = "hadley")
transform(head(hadleyverse), title = substr(title, 1, 50))
```

By using the result, you can install all packages created by hadley.

```{r eval=FALSE}
repos <- with(hadleyverse, paste(username, package_name, sep="/"))
githubinstall(repos) # I have not tried it
```

### 3.4. Search Packages by a Keyword

`gh_search_packages()` returns a list of R package repositories on GitHub that their titles contain a given keyword.

For example, if you want to search packages that are relevant to *lasso*, run the following.

```{r eval=FALSE}
gh_search_packages("lasso")
```

```{r echo=FALSE}
transform(gh_search_packages("lasso"), title = paste0(" ", substr(title, 1, 35), ".."))
```

### 3.5. Show the Source Code of Functions on GitHub

`gh_show_source()` looks for a source code on GitHub for a given function and tries to open the place on your Web browser.

```{r eval=FALSE}
gh_show_source("mutate", repo = "dplyr")
```

If you have loaded the package that the function belongs to, you can input the function directly.

```{r eval=FALSE}
library(dplyr)
gh_show_source(mutate)
```

This function may do not work well with Safari.

### 3.6. Update the List of R Packages

The **githubinstall** package uses [Gepuro Task Views](http://rpkg.gepuro.net) for getting the list of R packages on GitHub.
Gepuro Task Views is crawling the GitHub and updates information every day.
The package downloads the list of R packages from Gepuro Task Views each time it was loaded.
Thus, you can always use the newest list of packages on a new R session.

However, you may use an R session for a long time.
In such case, `gh_update_package_list()` is useful.

`gh_update_package_list()` updates the downloaded list of the R packages explicitly.

```{r eval=FALSE}
gh_update_package_list()
```

## 4. Related Work

- devtools: [Tools to make an R developer's life easier](https://github.com/hadley/devtools)
- ghit: [Lightweight GitHub Package Installer](https://github.com/cloudyr/ghit)
- drat: [Drat R Archive Template](https://github.com/eddelbuettel/drat)
- pacman: [A package management tools for R](https://github.com/trinker/pacman)
- remotes: [Install R packages from GitHub, Bitbucket, git, svn repositories, URLs](https://github.com/MangoTheCat/remotes)
